Crossing Origins by Crossing Formats

### 2013-02-06 main.tex
Some updates on the todo list

First steps in exploiting the formats in the context of the web have
been by researchers. Two noteworthy examples are GIFAR~\cite{Brandis:GIFAR} and
cross-origin CSS attacks~\cite{Huang+:CCS10} (Section~\ref{sec:related}
discusses further related work). 
%
GIFAR is based on polyglots that combine the GIF and JAR (Java
archive) formats. The former is used as benign and the latter as
malicious to bypass SOP.
%
Cross-origin CSS attacks inject fragments of CSS code into an existing
web page to extract information from the existing web page.
%
The paper is the first to perform a systematic study of polyglot
attacks.
We characterize of what constitutes a dangerous format in the web setting and
identify particularly dangerous formats, with PDF as the prime example. We
demonstrate that polyglot-based attacks on the web open up for insecure
communication across Internet origins. We perform a systematic study of
PDF-based injection and content smuggling attacks.

@@

Other examples of problematic elements include the \emph{object} and
\emph{embed} tags. These 
tags allow inclusion of resources that might require a 
plug-in to run. As with images, the content handled by the plug-in retains its origin, 
but the browsers are forced to rely on the plug-ins to employ correct security 
measures to prevent information leaks. The plug-in is selected based on the MIME-type of the content, but
because the server delivering the content might not be able to determine its
format, the \emph{object} tag allows a developer to set the type-attribute to guide the
browser in which plug-in to run. When the type attribute is used, the
corresponding plug-in is run regardless of which MIME-type the server returns
and it is up to the plug-in to respond to and handle the content it is served. While the containing page is restricted 
from directly accessing the content handled by the plug-in, a number of plugins provide 
an API for interacting with the plug-in and the data. As we will show, this can lead to 
to cross-origin information leaks.

@@

A polyglot is composed from various constructs that are either common between
languages; or constructs that are language specific, but carrying different
meaning in each language. To maintain validity across languages, one must ensure
that language specific constructs are not interpreted by other languages. This
is often accomplished by hiding the parts of the code specific to one language,
in parts interpreted as comments or plain text of another.

@@

Barth et al.~\cite{Barth+:SP09} define \emph{chameleon} documents, 
in which a benign file type crossbred with malicious HTML content. The document 
is crafted in such a way that the content-sniffing algorithm used in the browser 
will detect the content type as HTML. The paper proceeds to 
describe how an attacker can create a 
\emph{chameleon} document, that is valid PostScipt, but will be identified as HTML.
This issue allows exploitation when there is a mismatch between a websites 
content validation filter and the browsers content-sniffing algorithm.

In literature we find, apart from chameleons, other names for similar,
related concepts. Brandis~\cite{Brandis:GIFAR} refer \emph{GIFAR} 
attacks, based on one of the early instances of attacks based on 
GIF/JAR polyglots. 
Sundareswaran et al.~\cite{Sundareswaran:Squicciarini:CEAS10,Sundareswaran:Squicciarini:SecureComm10}
discuss GIFAR-related attacks as a form of \emph{content repurposing} attack. 
In this paper the term content smuggling is used as it best represent 
how a polyglot can infiltrate an origin.

\subsection{Scenario}

@@

Two attacker-centric components requires special attention, the infiltration of 
an origin and the exploitation of an infiltrated origin.
%In this situation, the same-origin policy is intended to prevent cross-origin 
%interaction.

\begin{figure}[h!]
  \center
  \includegraphics[width=0.49\textwidth]{scenario.png}
  \caption{Overview of the scenario}
  \label{fig:scenario}
\end{figure}

\subsection{Infiltrating an origin}

In the scenario the attacker infiltrate the origin by uploading a polyglot. Infiltration could, however, be performed in more than one way
This paper identifies two vectors for infiltrating an origin, via syntax 
injection, or content smuggling.

\paragraph{Syntax injection}

In a cross-site scripting attack, a HTML document is composed from input 
containing fragments of HTML syntax, causing the execution of attacker supplied 
script code. Similarly, in a syntax injection attack the vulnerable target will 
compose content from attacker controlled input containing syntax of a foreign 
format, thereby creating a polyglot. 
%
The target is assumed to employ server-side cross-site scripting sanitization. Hence, for the attack to be successful, the input must 
bypass this sanitization. The sanitization filter will remove or encode 
problematic characters, based on the context of the web application. In a polyglot attack
the injected format would pose as a new context, unknown to the filter. Chances are that the 
key elements of the injected format will pass the filter unaltered. 
The polyglot content can
be interpreted as either the benign original format or the
malicious attacker-supplied format. 
%

\paragraph{Content smuggling}
The vulnerable target of a polyglot attack based on content smuggling lets 
users to submit files to the service which are then served under the origin of 
the service. Examples of such services are file sharing services, image databases, 
social networks or job broker services. Such a service 
accepts a limited set of file-types, and the files are verified to
be of one of these types, e.g. an image, before being served to the end
user. By submitting a polyglot to such a service, an attacker can trick the
service into serving malicious content under the origin of the service. The
service will verify the file, assuming that it is of the benign format, but once
the content is served an attacker can coerce the content to be interpreted as
the malicious format.

\subsection{Exploiting the infiltrated origin}

@@

If the format also has the capability of communicating with the attacker, then 
the attacker can sensitive user information and leak it across origins. If the 
format is not restricted by the SOP, it can communicate directly with the 
attackers server. Otherwise, if the format can interact with the document that 
embeds the polyglot, the communication could be tunneled through this channel.

@@

\subsection{PDF-based polyglots}
Several design choices in the PDF specification make the format particularly 
suitable for making polyglots. One such design choice is the error tolerant 
parser. This is in part motivated it being a container format, of which the 
implication is that a PDF file can, by design, contain foreign syntax that 
could interfere with the syntax of the file. Another motivation is that PDF 
files are designed to be both forward- and backward-compatible. Readers 
implementing an older version of the specification do not recognize new 
features and behave as if they were not present in the file. 
%
The implementation notes of the specification describes some exceptions to the
requirements of the specification, such as the header can occur 
anywhere within the first 1024 bytes of the file. This flexibility gives plenty of room for 
combining with syntax of another format.
%
The specification declares many components to be required in a
PDF file, but as can be seen in Section~\ref{sec:format}, in practice several 
components can be omitted~\cite{Wolf:CCC10}. The code in Listing~\ref{lst:minimal-pdf} shows the 
minimal syntax required for a malformed, but valid PDF file.
%incrementally 
%updated by appending updated versions of existing objects to 
%the end of the file. In fact, 
%and formats. This and the fact that it is designed to be robust against errors,
%provides us with a format that is powerful, flexible and ideal for creating
%malicious polyglots.

@@

When a PDF document is embedded in a web page, the corresponding plugin 
is executed to render the content. Recall from Section~\ref{sec:cross-origin} 
that the plug-in is selected based 
on the MIME-type, either supplied by the server or in the type attribute, and 
that the browser rely on the plug-in to handle the situation when the MIME-type 
supplied by the server is in appropriate for the plug-in.
In the case of the Adobe Reader plug-in, it disregards the MIME-type
supplied by the server and will attempt to interpret any content as PDF.

@@

\paragraph{HTML comments}
No HTML enclosed in comments, "<--" and "-->", will be rendered, and therefore 
filters consider this context safe. To prevent input from escaping the comment by 
injecting an end comment token, certain filters remove any occurrences of "-->", 
but leave the of the input untouched. HTML comments are meaningless to PDF, 
and the result is a valid HTML/PDF polyglot.

\paragraph{JavaScript strings}
In an in-line JavaScript context, user input is often included in the shape of a 
JavaScript string, delimited by single or double quotes. In this context only a 
few characters require encoding, as they can break the string context. Naturally, 
any single or double quotes need escaping, as well as any carriage return or 
line feed characters. In the spirit of making minimal changes to the input, 
certain websites only these characters. This is sufficient to prevent cross-site 
scripting attacks, but fail to protect against PDF-based polyglot attacks.

@@

Due to the nature of the PDF, it can without much effort be combined with just 
about any other format. This provides ample opportunity to create malicious 
polyglots where PDF is either the benign or the malicious format.

@@

To better understand the how this problem presents itself, a comparison of browser and reader instances 
is presented. We compare all major browsers and two of the most common readers.

@@

Section~\ref{sec:vuln} focus on the Adobe PDF Reader as the attack surface, 
due to its standing as the most commonly used reader. To give a comparison, the 
Google Chrome built-in PDF reader, was selected as it is the default reader to 
user of the browser.

@@

The current best-practice recommendation from Google~\cite{Google:ContentHosting} 
on hosting user-supplied content, is to serve the content from an origin that is 
completely separate, and isolated from the sensitive services. It also describes how 
Google moved away from sanitizing content, as it was found to be insufficient in 
many cases.

Google uses the domain name \emph{googleusercontent.com} to sandbox all the 
untrusted user files that their various services provide for their customers.

>>>>

The first steps in exploiting formats in the context of the web have
been taken by researchers. Two noteworthy examples are GIFAR~\cite{Brandis:GIFAR} and
cross-origin CSS attacks~\cite{Huang+:CCS10} (Section~\ref{sec:related}
discusses further related work). 
%
GIFAR is based on polyglots that combine the GIF and JAR (Java
archive) formats. The former is used as benign and the latter as
malicious to bypass SOP.
%
Cross-origin CSS attacks inject fragments of CSS code into an existing
web page to extract information from the existing web page.
%
The paper is the first to perform a systematic study of polyglot
attacks.
We characterize of what constitutes a dangerous format in the web setting and
identify particularly dangerous formats, with PDF as the prime example. We
demonstrate that polyglot-based attacks on the web open up for insecure
communication across Internet origins. We perform a systematic study of
PDF-based injection and content smuggling attacks.

@@

Other examples of problematic elements include the \emph{object} and
\emph{embed} tags. These 
tags allow inclusion of resources that might require a 
plug-in to run. As with images, the content handled by the plug-in retains its origin, 
but the browsers are forced to rely on the plug-ins to employ correct security 
measures to prevent information leaks. The plug-in is selected based on the MIME-type of the content, but
because the server delivering the content might not be able to determine its
format, the \emph{object} tag allows a developer to set the type-attribute to guide the
browser in which plug-in to run. When the type attribute is used, the
corresponding plug-in is run regardless of which MIME-type the server returns
and it is up to the plug-in to respond to and handle the content it is served. While the containing page is restricted 
from directly accessing the content handled by the plug-in, a number of plug-ins provide 
an API for interacting with the plug-in and the data. As we will show, this can lead to 
to cross-origin information leaks.

@@

A polyglot is composed from various constructs that are either common between
languages; or constructs that are language specific but carrying different
meaning in each language. To maintain validity across languages, one must ensure
that language specific constructs are not interpreted by other languages. This
is often accomplished by hiding the parts of the code specific to one language,
in parts interpreted as comments or plain text of another.

@@

Barth et al.~\cite{Barth+:SP09} define \emph{chameleon} documents, 
in which a benign file type crossbred with malicious HTML content. The document 
is crafted in such a way that the content-sniffing algorithm used in the browser 
will detect the content type as HTML. The paper proceeds to 
describe how an attacker can create a 
\emph{chameleon} document, that is valid PostScript, but will be identified as HTML.
This issue allows exploitation when there is a mismatch between a websites 
content validation filter and the browsers content-sniffing algorithm.

In literature we find, apart from chameleons, other names for similar,
related concepts. Brandis~\cite{Brandis:GIFAR} refer to \emph{GIFAR} 
attacks, based on one of the early instances of attacks based on 
GIF/JAR polyglots. 
Sundareswaran et al.~\cite{Sundareswaran:Squicciarini:CEAS10,Sundareswaran:Squicciarini:SecureComm10}
discuss GIFAR-related attacks as a form of \emph{content repurposing} attack. 
In this paper the term content smuggling is used as it best represents 
how a polyglot can infiltrate an origin.

Two attacker-centric components require special attention, the infiltration of 
an origin and the exploitation of an infiltrated origin.
%In this situation, the same-origin policy is intended to prevent cross-origin 
%interaction.

\subsection{Attack vectors}

In the scenario the attacker infiltrate the origin by uploading a polyglot. Infiltration could, however, be performed in more than one way.
This paper identifies two vectors for infiltrating an origin, via syntax 
injection, or content smuggling.

In a cross-site scripting attack, an HTML document is composed from input 
containing fragments of HTML syntax, causing the execution of attacker supplied 
script code. Similarly, in a syntax injection attack the vulnerable target will 
compose content from attacker controlled input containing syntax of a foreign 
format, thereby creating a polyglot. 
%
The target is assumed to employ server-side cross-site scripting sanitization. Hence, for the attack to be successful, the input must 
bypass this sanitization. The sanitization filter will remove or encode 
problematic characters, based on the context of the web application. In a polyglot attack
the injected format would pose as a new context, unknown to the filter. Chances are that the 
key elements of the injected format will pass the filter unaltered. 
The polyglot content can
be interpreted as either the benign original format or the
malicious attacker-supplied format. 
%

\begin{figure}[h!]
  \center
  \includegraphics[width=0.49\textwidth]{injection-scenario.png}
  \caption{Overview of the syntax injection scenario}
  \label{fig:scenario}
\end{figure}

\subsubsection{Content smuggling}
The vulnerable target of a polyglot attack, based on content smuggling, lets 
users to submit files to the service which are subsequently served under the origin of 
the service. Examples of such services are file sharing services, image databases, 
social networks or job broker services. Such a service 
accepts a limited set of file-types, and the files are verified to
be of one of these types, e.g. an image, before being served to the end
user. By submitting a polyglot to such a service, an attacker can trick the
service into serving malicious content under the origin of the service. The
service will verify the file, assuming that it is of the benign format, but once
the content is served an attacker can coerce the content to be interpreted as
the malicious format.

\paragraph{Scenario}

@@

\begin{figure}[h!]
  \center
  \includegraphics[width=0.49\textwidth]{smuggling-scenario.png}
  \caption{Overview of the content smuggling scenario}
  \label{fig:scenario}
\end{figure}

\subsection{Payloads}

@@

If the format also has the capability of communicating with the attacker, then 
the attacker can extract sensitive user information and leak it across origins. If the 
format is not restricted by the SOP, it can communicate directly with the 
attackers server. Otherwise, if the format can interact with the document that 
embeds the polyglot, the communication could be tunneled through this channel.

@@

\subsection{PDF-based polyglots}
Several design choices in the PDF specification make the format particularly 
suitable for making polyglots. One such design choice is the error tolerant 
parser. This is in part motivated by PDF being a container format, of which the 
implication is that a PDF file can, by design, contain foreign syntax that 
could interfere with the syntax of the file. Another motivation is that PDF 
files are designed to be both forward- and backward-compatible. Readers 
implementing an older version of the specification do not recognize new 
features and behave as if they were not present in the file. 
%
The implementation notes of the specification describes some exceptions to the
requirements of the specification, such as the header can occur 
anywhere within the first 1024 bytes of the file. This flexibility gives plenty of room for 
combining with syntax of another format.
%
The specification declares many components to be required in a
PDF file, but as can be seen in Section~\ref{sec:format}, in practice several 
components can be omitted~\cite{Wolf:CCC10}. The code in Listing~\ref{lst:minimal-pdf} shows the 
minimal syntax required for a malformed, but valid PDF file.
%incrementally 
%updated by appending updated versions of existing objects to 
%the end of the file. In fact, 
%and formats. This and the fact that it is designed to be robust against errors,
%provides us with a format that is powerful, flexible and ideal for creating
%malicious polyglots.

@@

When a PDF document is embedded in a web page, the corresponding plugin 
is executed to render the content. Recall from Section~\ref{sec:cross-origin} 
that the plug-in is selected based 
on the MIME-type, either supplied by the server or in the type attribute, and 
that the browser rely on the plug-in to handle the situation when the MIME-type 
supplied by the server is inappropriate for the plug-in.
In the case of the Adobe Reader plug-in, it disregards the MIME-type
supplied by the server and will attempt to interpret any content as PDF.

@@

\paragraph{HTML comments}
No HTML enclosed in comments, "<--" and "-->", will be rendered, and therefore 
filters consider this context safe. To prevent input from escaping the comment by 
injecting an end comment token, certain filters remove any occurrences of "-->", 
but leave the rest of the input untouched. HTML comments are meaningless to PDF, 
and the result is a valid HTML/PDF polyglot.

\paragraph{JavaScript strings}
In an in-line JavaScript context, user input is often included in the shape of a 
JavaScript string, delimited by single or double quotes. In this context only a 
few characters require encoding, as they can break the string context. Naturally, 
any single or double quotes need escaping, as well as any carriage return or 
line feed characters. In the spirit of making minimal changes to the input, 
certain websites only escape these characters. This is sufficient to prevent cross-site 
scripting attacks, but fail to protect against PDF-based polyglot attacks.

@@

Due to the nature of the PDF, it can without much effort be combined with just 
about any other format. This provides ample opportunity to create malicious 
polyglots where PDF is either the benign or the malicious format. Consequently, 
this significantly expands the attack surface, making it important to take 
measures to protect against these attacks.

@@

To better understand how this problem presents itself, a comparison of browser and reader instances 
is presented. We compare all major browsers and two of the most common readers.

@@

Section~\ref{sec:vuln} focuses on the Adobe PDF Reader as the attack surface, 
due to its standing as the most commonly used reader. To give a comparison, the 
Google Chrome built-in PDF reader, was selected as it is the default reader to 
user of the browser.

@@

The current best-practice recommendation from Google~\cite{Google:ContentHosting} 
on hosting user-supplied content, is to serve the content from an origin that is 
completely separate, and isolated from the sensitive services. It also describes how 
Google moved away from sanitizing content, as it was found to be insufficient in 
many cases. These best-practices successfully prevents content-smuggling 
attacks in all aspects, as the infiltrated origin is separate from the origin 
containing sensitive data.

%Google uses the domain name \emph{googleusercontent.com} to sandbox all the 
%untrusted user files that their various services provide for their customers.
